# The Tidy Text Format {#tidytext}

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE)

options(width = 100, dplyr.width = 100)
```

Intro text may go here about the one-token-per-document-per-row and about what is explored in the chapter.

## The `unnest_tokens` function

```{r text}
text <- c("Because I could not stop for Death -",
          "He kindly stopped for me -",
          "The Carriage held but just Ourselves -",
          "and Immortality")

text
```

This is a typical character vector that we might want to analyze. In order to turn it into a tidy text dataset, we first need to put it into a data frame:

```{r text_df, dependson = "text"}
library(dplyr)
text_df <- data_frame(line = 1:4, text = text)

text_df
```

Notice that this data frame isn't yet compatible with tidy tools. We can't filter out words or count which occur most frequently, since each row is made up of multiple coimbined tokens. We need to turn this into **one-token-per-document-per-row**.

To do this, we use tidytext's `unnest_tokens` function:

```{r dependson = "text_df", R.options = list(dplyr.print_max = 10)}
library(tidytext)
text_df %>%
  unnest_tokens(word, text)
```

We've now split each row so that there's one token (word) in each row of the new data frame. Also notice:

* Other columns, such as the line number each word came from, are retained
* Punctuation has been stripped
* By default, `unnest_tokens` turns the tokens lowercase, which makes them easier to compare or combine with other datasets. (Use the `to_lower = FALSE` argument to turn off this behavior).

Having the text data in this format lets us manipulate, process, and visualize the text using the standard set of tidy tools; namely dplyr, tidyr, ggplot2, and broom.

## Example: the works of Jane Austen

Let's use the text of Jane Austen's 6 completed, published novels from the [janeaustenr](https://cran.r-project.org/package=janeaustenr) package, and transform them into a tidy format. janeaustenr provides them as a one-row-per-line format:

```{r original_books}
library(janeaustenr)
library(dplyr)
library(stringr)

original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]",
                                                 ignore_case = TRUE)))) %>%
  ungroup()

original_books
```

To work with this as a tidy dataset, we need to restructure it as **one-token-per-row** format. The `unnest_tokens` function is a way to convert a dataframe with a text column to be one-token-per-row:

```{r tidy_books_raw, dependson = "original_books"}
library(tidytext)
tidy_books <- original_books %>%
  unnest_tokens(word, text)

tidy_books
```

This function uses the [tokenizers package](https://github.com/lmullen/tokenizers) to separate each line into words. The default tokenizing is for words, but other options include characters, ngrams, sentences, lines, paragraphs, or separation around a regex pattern.

Now that the data is in one-word-per-row format, we can manipulate it with tidy tools like dplyr. We can remove stop words (kept in the tidytext dataset `stop_words`) with an `anti_join`.

```{r tidy_books, dependson = "tidy_books_raw"}
data("stop_words")

tidy_books <- tidy_books %>%
  anti_join(stop_words)
```

We can also use `count` to find the most common words in all the books as a whole.

```{r dependson = "tidy_books"}
tidy_books %>%
  count(word, sort = TRUE) 
```

For example, this allows us to visualize the popular words using ggplot2:

```{r}
library(ggplot2)

tidy_books %>%
  count(word, sort = TRUE)
```

## The gutenbergr package

TODO: Now that we've introduced the janeaustenr package, also include a brief intro to the gutenberg package.

## Word Frequencies

A common task in text mining is to look at word frequencies and to compare frequencies across different texts. We can do this intuitively and smoothly using tidy data principles. We already have Jane Austen's works; let's get two more sets of texts to compare to. First, let's look at some science fiction and fantasy novels by H.G. Wells, who lived in the late 19th and early 20th centuries. Let's get [*The Time Machine*](https://www.gutenberg.org/ebooks/35), [*The War of the Worlds*](https://www.gutenberg.org/ebooks/36), [*The Invisible Man*](https://www.gutenberg.org/ebooks/5230), and [*The Island of Doctor Moreau*](https://www.gutenberg.org/ebooks/159).

```{r}
library(gutenbergr)
hgwells <- gutenberg_download(c(35, 36, 5230, 159))
tidy_hgwells <- hgwells %>%
        unnest_tokens(word, text) %>%
        anti_join(stop_words)
```

Just for kicks, what are the most common words in these novels of H.G. Wells?

```{r}
tidy_hgwells %>%
        count(word, sort = TRUE)
```

Now let's get some well-known works of the Brontë sisters, whose lives overlapped with Jane Austen's somewhat but who wrote in a bit of a different style. Let's get [*Jane Eyre*](https://www.gutenberg.org/ebooks/1260), [*Wuthering Heights*](https://www.gutenberg.org/ebooks/768), [*The Tenant of Wildfell Hall*](https://www.gutenberg.org/ebooks/969), [*Villette*](https://www.gutenberg.org/ebooks/9182), and [*Agnes Grey*](https://www.gutenberg.org/ebooks/767).

```{r}
bronte <- gutenberg_download(c(1260, 768, 969, 9182, 766))
tidy_bronte <- bronte %>%
        unnest_tokens(word, text) %>%
        anti_join(stop_words)
```

What are the most common words in these novels of the Brontë sisters?

```{r}
tidy_bronte %>%
        count(word, sort = TRUE)
```

Well, Jane Austen is not going around talking about people's *hearts* this much; I can tell you that right now. Those Brontë sisters, SO DRAMATIC. Interesting that "time" and "door" are in the top 10 for both H.G. Wells and the Brontë sisters. "Door"?!

Anyway, let's calculate the frequency for each word for the works of Jane Austen, the Brontë sisters, and H.G. Wells.

```{r}
tidy_both <- bind_rows(
        mutate(tidy_bronte, author = "Brontë Sisters"),
        mutate(tidy_hgwells, author = "H.G. Wells"))
frequency <- tidy_both %>%
        mutate(word = str_extract(word, "[a-z]+")) %>%
        count(author, word) %>%
        rename(other = n) %>%
        inner_join(count(tidy_books, word)) %>%
        rename(Austen = n) %>%
        mutate(other = other / sum(other),
               Austen = Austen / sum(Austen)) %>%
        ungroup()
```

I'm using `str_extract` here because the UTF-8 encoded texts from Project Gutenberg have some examples of words with underscores around them to indicate emphasis (like italics). The tokenizer treated these as words but we don't want to count "\_any\_" separately from "any". Now let's plot.

```{r, fig.width=10, fig.height=5.5}
library(scales)
ggplot(frequency, aes(x = other, y = Austen, color = abs(Austen - other))) +
        geom_abline(color = "gray40") +
        geom_jitter(alpha = 0.1, size = 2.5, width = 0.4, height = 0.4) +
        geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
        scale_x_log10(labels = percent_format()) +
        scale_y_log10(labels = percent_format()) +
        scale_color_gradient(limits = c(0, 0.001), low = "darkslategray4", high = "gray75") +
        facet_wrap(~author, ncol = 2) +
        theme_minimal(base_size = 14) +
        theme(legend.position="none") +
        labs(title = "Comparing Word Frequencies",
             subtitle = "Word frequencies in Jane Austen's texts are closer to the Brontë sisters' than to H.G. Wells'",
             y = "Jane Austen", x = NULL)
```

Words that are close to the line in these plots have similar frequencies in both sets of texts, for example, in both Austen and Brontë texts ("miss", "time", "lady", "day" at the upper frequency end) or in both Austen and Wells texts ("time", "day", "mind", "brother" at the high frequency end). Words that are far from the line are words that are found more in one set of texts than another. For example, in the Austen-Brontë plot, words like "elizabeth", "emma", "captain", and "bath" (all proper nouns) are found in Austen's texts but not much in the Brontë texts, while words like "arthur", "dark", "dog", and "doctor" are found in the Brontë texts but not the Austen texts. In comparing H.G. Wells with Jane Austen, Wells uses words like "beast", "guns", "brute", and "animal" that Austen does not, while Austen uses words like "family", "friend", "letter", and "agreeable" that Wells does not.

Overall, notice that the words in the Austen-Brontë plot are closer to the zero-slope line than in the Austen-Wells plot and also extend to lower frequencies; Austen and the Brontë sisters use more similar words than Austen and H.G. Wells. Also, we might notice the percent frequencies for individual words are different in one plot when compared to another because of the inner join; not all the words are found in all three sets of texts so the percent frequency is a different quantity.

Let's quantify how similar and different these sets of word frequencies are using a correlation test. How correlated are the word frequencies between Austen and the Brontë sisters, and between Austen and Wells?

```{r}
cor.test(data = frequency[frequency$author == "Brontë Sisters",], ~ other + Austen)
cor.test(data = frequency[frequency$author == "H.G. Wells",], ~ other + Austen)
```

The relationship between the word frequencies is different between these sets of texts, as it appears in the plots.
