# Sentiment analysis with tidy data {#sentiment}

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE)
options(width = 100, dplyr.width = 100)
library(ggplot2)
theme_set(theme_light())
```

## The `sentiments` dataset

There are a variety of methods and dictionaries that exist for evaluating the opinion or emotion in text. The tidytext package contains three sentiment lexicons in the `sentiments` dataset. 

```{r}
library(tidytext)

sentiments
```

The three lexicons are

* `AFINN` from [Finn Årup Nielsen](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010),
* `bing` from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html), and
* `nrc` from [Saif Mohammad and Peter Turney](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm).

All three of these lexicons are based on unigrams (or single words). These lexicons contain many English words and the words are assigned scores for positive/negative sentiment, and also possibly emotions like joy, anger, sadness, and so forth. The `nrc` lexicon categorizes words in a binary fashion ("yes"/"no") into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust. The `bing` lexicon categorizes words in a binary fashion into positive and negative categories. The `AFINN` lexicon assigns words with a score that runs between -5 and 5, with negative scores indicating negative sentiment and positive scores indicating positive sentiment. All of this information is tabulated in the `sentiments` dataset.

These dictionary-based methods find the total sentiment of a piece of text by adding up the individual sentiment scores for each word in the text. Not every English word is in the lexicons because many English words are pretty neutral. It is important to keep in mind that these methods do not take into account qualifiers before a word, such as in "no good" or "not true"; a lexicon-based method like this is based on unigrams only. For many kinds of text (like the narrative examples below), there are not sustained sections of sarcasm or negated text, so this is not an important effect. 

One last caveat is that the size of the chunk of text that we add up unigram sentiment scores for can have an important effect for an analysis. A paragraph-sized text can often have positive and negative sentiment averaged out to about zero, while sentence-sized text often works better.

## Sentiment analysis with inner join

With data in a tidy format, sentiment analysis can be done as an inner join. Let's look at the words with a joy score from the NRC lexicon. What are the most common joy words in *Emma*?

```{r tidy_books}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r nrcjoy, dependson = "tidy_books"}
nrcjoy <- sentiments %>%
  filter(lexicon == "nrc", sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  semi_join(nrcjoy) %>%
  count(word, sort = TRUE)
```

Or instead we could examine how sentiment changes during each novel. Let's find a sentiment score for each word using the Bing lexicon, then count the number of positive and negative words in defined sections of each novel.

```{r janeaustensentiment, dependson = "tidy_books"}
library(tidyr)
bing <- sentiments %>%
  filter(lexicon == "bing") %>%
  select(-score)

janeaustensentiment <- tidy_books %>%
  inner_join(bing) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

Now we can plot these sentiment scores across the plot trajectory of each novel.

```{r sentimentplot, dependson = "janeaustensentiment", fig.width=9, fig.height=10}
library(ggplot2)

ggplot(janeaustensentiment, aes(index, sentiment, fill = book)) +
        geom_bar(alpha = 0.8, stat = "identity", show.legend = FALSE) +
        facet_wrap(~book, ncol = 2, scales = "free_x")
```

We can see here how the plot of each novel changes toward more positive or negative sentiment over the trajectory of the story.

TODO: The three different methods of calculating sentiment give results that are different in an absolute sense but have similar relative trajectories through the novel.

## Most common positive and negative words

One advantage of having the data frame with both sentiment and word is that we can analyze word counts that contribute to each sentiment.

```{r wordcounts, fig.width=7, fig.height=6, dependson = "tidy_books"}
bing_word_counts <- tidy_books %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

This can be shown visually, and we can pipe straight into ggplot2, if we like, because of the way we are consistently using tools built for handling tidy data frames.

```{r pipetoplot, dependson = "wordcounts", fig.width=6, fig.height=6}
bing_word_counts %>%
  filter(n > 150) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_bar(alpha = 0.8, stat = "identity") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()
```

This lets us spot an anomaly in the sentiment analysis; the word "miss" is coded as negative but it is used as a title for young, unmarried women in Jane Austen's works. If it were appropriate for our purposes, we could easily add "miss" to a custom stop-words list using `bind_rows`.

## Wordclouds

We've seen that this tidy text mining approach works well with ggplot2, but having our data in a tidy format is useful for other plots as well.

For example, consider the wordcloud package. Let's look at the most common words in Jane Austen's works as a whole again.

```{r firstwordcloud, dependson = "tidy_books", fig.height=6, fig.width=6}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

In other functions, such as `comparison.cloud`, you may need to turn the data frame into a matrix with reshape2's `acast`. Let's do the sentiment analysis to tag positive and negative words using an inner join, then find the most common positive and negative words. Until the step where we need to send the data to `comparison.cloud`, this can all be done with joins, piping, and dplyr because our data is in tidy format.

```{r wordcloud, dependson = "tidy_books", fig.height=5, fig.width=5}
library(reshape2)

tidy_books %>%
  inner_join(bing) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("#F8766D", "#00BFC4"),
                   max.words = 100)
```

## Looking at units beyond just words

Lots of useful work can be done by tokenizing at the word level, but sometimes it is useful or necessary to look at different units of text. For example, some sentiment analysis algorithms look beyond only unigrams (i.e. single words) to try to understand the sentiment of a sentence as a whole. These algorithms try to understand that

> I am not having a good day.

is a sad sentence, not a happy one, because of negation. The [Stanford CoreNLP](http://stanfordnlp.github.io/CoreNLP/) tools and the [sentimentr R package](https://github.com/trinker/sentimentr) (currently available on Github but not CRAN) are examples of such sentiment analysis algorithms. For these, we may want to tokenize text into sentences.

```{r PandP}
PandP_sentences <- data_frame(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
```

Let's look at just one.

```{r PandPsentences, dependson = "PandP"}
PandP_sentences$sentence[2]
```

The sentence tokenizing does seem to have a bit of trouble with UTF-8 encoded text, especially with sections of dialogue; it does much better with punctuation in ASCII. One possibility, if this is important, is to try using `iconv()`, with something like `iconv(text, to = 'latin1')` in a mutate statement before unnesting.

Another option in `unnest_tokens` is to split into tokens using a regex pattern. We could use this, for example, to split the text of Jane Austen's novels into a data frame by chapter.

```{r austen_chapters, dependson = "tidy_books"}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

We have recovered the correct number of chapters in each novel (plus an "extra" row for each novel title). In this data frame, each row corresponds to one chapter.

Near the beginning of this vignette, we used a similar regex to find where all the chapters were in Austen's novels for a tidy data frame organized by one-word-per-row. We can use tidy text analysis to ask questions such as what are the most negative chapters in each of Jane Austen's novels? First, let's get the list of negative words from the Bing lexicon. Second, let's make a dataframe of how many words are in each chapter so we can normalize for the length of chapters. Then, let's find the number of negative words in each chapter and divide by the total words in each chapter. Which chapter has the highest proportion of negative words?

```{r chapters, dependson = "tidy_books"}
bingnegative <- sentiments %>%
  filter(lexicon == "bing", sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup
```

These are the chapters with the most sad words in each book, normalized for number of words in the chapter. What is happening in these chapters? In Chapter 43 of *Sense and Sensibility* Marianne is seriously ill, near death, and in Chapter 34 of *Pride and Prejudice* Mr. Darcy proposes for the first time (so badly!). Chapter 46 of *Mansfield Park* is almost the end, when everyone learns of Henry's scandalous adultery, Chapter 15 of *Emma* is when horrifying Mr. Elton proposes, and in Chapter 21 of *Northanger Abbey* Catherine is deep in her Gothic faux fantasy of murder, etc. Chapter 4 of *Persuasion* is when the reader gets the full flashback of Anne refusing Captain Wentworth and how sad she was and what a terrible mistake she realized it to be.
