# Tidying and casting document-term matrices

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE)

options(width = 100, dplyr.width = 150)
```

Intro text here.

## Tidying a document-term matrix

Many existing text mining datasets are in the form of a `DocumentTermMatrix` class (from the tm package). For example, consider the corpus of 2246 Associated Press articles from the topicmodels package:

```{r AssociatedPress}
library(tm)
data("AssociatedPress", package = "topicmodels")
AssociatedPress
```

If we want to analyze this with tidy tools, we need to turn it into a one-token-per-document-per-row data frame first. The `tidy` function does this. (For more on the tidy verb, [see the broom package](https://github.com/dgrtwo/broom)).

```{r ap_td, dependson = "AssociatedPress"}
library(dplyr)
library(tidytext)

ap_td <- tidy(AssociatedPress)
```

Just as shown in [this vignette](tidytext.html), having the text in this format is convenient for analysis with the tidytext package. For example, you can perform sentiment analysis on these newspaper articles.

```{r ap_sentiments, dependson = "ap_td"}
bing <- sentiments %>%
  filter(lexicon == "bing") %>%
  select(word, sentiment)

ap_sentiments <- ap_td %>%
  inner_join(bing, by = c(term = "word"))

ap_sentiments
```

We can find the most negative documents:

```{r most_negative_docs, dependson = "ap_sentiments"}
library(tidyr)

ap_sentiments %>%
  count(document, sentiment, wt = count) %>%
  ungroup() %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative) %>%
  arrange(sentiment)
```

Or visualize which words contributed to positive and negative sentiment:

```{r dependson = "ap_sentiments", fig.width = 7, fig.height = 5}
library(ggplot2)

ap_sentiments %>%
  count(sentiment, term, wt = count) %>%
  ungroup() %>%
  filter(n >= 150) %>%
  mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
  mutate(term = reorder(term, n)) %>%
  ggplot(aes(term, n, fill = sentiment)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylab("Contribution to sentiment")
```

Note that a tidier is also available for the `dfm` class from the quanteda package:

```{r quanteda_corpus}
data("inaugCorpus", package = "quanteda")
d <- quanteda::dfm(inaugCorpus)

d

tidy(d)
```

## Casting tidy text data into a DocumentTermMatrix

Some existing text mining tools or algorithms work only on sparse document-term matrices. Therefore, tidytext provides `cast_` verbs for converting from a tidy form to these matrices.

```{r}
ap_td

# cast into a Document-Term Matrix
ap_td %>%
  cast_dtm(document, term, count)

# cast into a Term-Document Matrix
ap_td %>%
  cast_tdm(term, document, count)

# cast into quanteda's dfm
ap_td %>%
  cast_dfm(term, document, count)


# cast into a Matrix object
m <- ap_td %>%
  cast_sparse(document, term, count)
class(m)
dim(m)
```

This allows for easy reading, filtering, and processing to be done using dplyr and other tidy tools, after which the data can be converted into a document-term matrix for machine learning applications.

## Tidying corpus objects with metadata

You can also tidy Corpus objects from the tm package. For example, consider a Corpus containing 20 documents:

```{r reuters}
reut21578 <- system.file("texts", "crude", package = "tm")
reuters <- VCorpus(DirSource(reut21578),
                   readerControl = list(reader = readReut21578XMLasPlain))

reuters
```

The `tidy` verb creates a table with one row per document:

```{r reuters_td, dependson = "reuters"}
reuters_td <- tidy(reuters)
reuters_td
```

Similarly, you can `tidy` a `corpus` object from the quanteda package:

```{r inaug_td}
library(quanteda)

data("inaugCorpus")

inaugCorpus

inaug_td <- tidy(inaugCorpus)
inaug_td
```

This lets us work with tidy tools like `unnest_tokens` to analyze the text alongside the metadata.

```{r inaug_words, dependson = "inaug_td"}
inaug_words <- inaug_td %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

inaug_words
```

We could then, for example, see how the appearance of a word changes over time:

```{r inaug_freq, dependson = "inaug_words"}
inaug_freq <- inaug_words %>%
  count(Year, word) %>%
  ungroup() %>%
  complete(Year, word, fill = list(n = 0)) %>%
  group_by(Year) %>%
  mutate(year_total = sum(n),
         percent = n / year_total) %>%
  ungroup()

inaug_freq
```

For example, we can use the broom package to perform logistic regression on each word.

```{r models, dependson = "inaug_freq"}
models <- inaug_freq %>%
  group_by(word) %>%
  filter(sum(n) > 50) %>%
  do(tidy(glm(cbind(n, year_total - n) ~ Year, .,
              family = "binomial"))) %>%
  ungroup() %>%
  filter(term == "Year")

models

models %>%
  filter(term == "Year") %>%
  arrange(desc(abs(estimate)))
```

You can show these models as a volcano plot, which compares the effect size with the significance:

```{r dependson = "models"}
library(ggplot2)
theme_set(theme_bw())

models %>%
  mutate(adjusted.p.value = p.adjust(p.value)) %>%
  ggplot(aes(estimate, adjusted.p.value)) +
  geom_point() +
  scale_y_log10() +
  geom_text(aes(label = word), vjust = 1, hjust = 1,
            check_overlap = TRUE) +
  xlab("Estimated change over time") +
  ylab("Adjusted p-value")
```

We can also use the ggplot2 package to display the top 6 terms that have changed in frequency over time.

```{r dependson = "models"}
library(scales)

models %>%
  top_n(6, abs(estimate)) %>%
  inner_join(inaug_freq) %>%
  ggplot(aes(Year, percent)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ word) +
  scale_y_continuous(labels = percent_format()) +
  ylab("Frequency of word in speech")
```
