# Case study: analyzing usenet text {#usenet}

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE)
options(width = 100, dplyr.width = 150)
library(ggplot2)
theme_set(theme_light())
```

Here we'll use what we've learned in the book to perform a start-to-finish analysis of a set of 20,000 messages sent to 20 Usenet bulletin boards in 1993. The Usenet bulletin boards in this data set include boards for topics like politics, autos, "for sale", atheism, etc. This data set is [publicly available](http://qwone.com/~jason/20Newsgroups/) and has become popular for testing and exercises in text analysis and machine learning.

## Setup

We'll start by reading in all the messages. (Note that this step takes several minutes).

```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
```

```{r eval = FALSE}
training_folder <- "data/20news-bydate/20news-bydate-train/"

read_folder <- function(infolder) {
  print(infolder)
  data_frame(file = dir(infolder, full.names = TRUE)) %>%
    mutate(text = map(file, read_lines)) %>%
    transmute(id = basename(file), text) %>%
    unnest(text)
}

raw_text <- data_frame(folder = dir(training_folder, full.names = TRUE)) %>%
  unnest(map(folder, read_folder)) %>%
  transmute(board = basename(folder), id, text)
```

```{r raw_text, echo = FALSE}
load("data/raw_text.rda")
# save(raw_text, file = "data/raw_text.rda")
```

Each email has structure we need to remove. For starters:

* Every email has one or more headers (e.g. "from:", "in_reply_to:")
* Many have signatures, which (since they're constant for each user) we wouldn't want to examine alongside the content

We need to remove headers and signatures:

```{r dependson = "raw_text"}
# remove headers and signatures
cleaned_text <- raw_text %>%
  group_by(id) %>%
  filter(cumsum(text == "") > 0,
         cumsum(str_detect(text, "^--")) == 0) %>%
  ungroup()

# remove nested text (starting with ">") and lines that note the author
# of those
cleaned_text <- cleaned_text %>%
  filter(str_detect(text, "^[^>]+[A-Za-z\\d]") | text == "",
         !str_detect(text, "writes(:|\\.\\.\\.)$"),
         !str_detect(text, "^In article <"),
         !id %in% c(9704, 9985))
```

```{r cleaned_text}
library(tidytext)

usenet_words <- cleaned_text %>%
  unnest_tokens(word, text) %>%
  filter(str_detect(word, "^[a-z]"),
         str_detect(word, "[a-z]$"),
         !word %in% stop_words$word)
```

We could simply find the most common words:

```{r}
usenet_words %>%
  count(word, sort = TRUE)
```

Or we could look at the most common words by board:

```{r words_by_board}
words_by_board <- usenet_words %>%
  count(board, word) %>%
  ungroup()
```

```{r dependson = "words_by_board"}
words_by_board %>%
  group_by(board) %>%
  top_n(3)
```

## Term frequency and inverse document frequency (tf-idf)

We notice that some words are likely to be more common on particular boards. Let's try quantifying this using the tf-idf metric we learned in [Chapter 4](tfidf).

```{r}
tf_idf <- words_by_board %>%
  bind_tf_idf(word, board, n) %>%
  arrange(desc(tf_idf))

tf_idf
```

We can visualize this for a few select boards. For example, let's look at all the `sci.` boards:

```{r, fig.width=9, fig.height=8}
library(ggplot2)
library(ggstance)

tf_idf %>%
  filter(str_detect(board, "^sci\\.")) %>%
  group_by(board) %>%
  top_n(12, tf_idf) %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word)) +
  geom_barh(stat = "identity") +
  facet_wrap(~ board, scales = "free_y") +
  xlab("tf-idf")
```

We could use almost the same code (not shown) to compare the "rec." (recreation) or "talk." boards:

```{r echo = FALSE, fig.width=9, fig.height=8}
plot_tf_idf <- function(d) {
  d %>%
    group_by(board) %>%
    top_n(10, tf_idf) %>%
    mutate(word = reorder(word, tf_idf)) %>%
    ggplot(aes(tf_idf, word)) +
    geom_barh(stat = "identity") +
    facet_wrap(~ board, scales = "free_y") +
    xlab("tf-idf")
}

tf_idf %>%
  filter(str_detect(board, "^rec\\.")) %>%
  plot_tf_idf()

tf_idf %>%
  filter(str_detect(board, "^talk\\.")) %>%
  plot_tf_idf()
```

## Sentiment Analysis

```{r}
AFINN <- sentiments %>%
  filter(lexicon == "AFINN")

word_board_sentiments <- words_by_board %>%
  inner_join(AFINN, by = "word")

board_sentiments <- word_board_sentiments %>%
  group_by(board) %>%
  summarize(score = sum(score * n) / sum(n))

board_sentiments %>%
  mutate(board = reorder(board, score)) %>%
  ggplot(aes(board, score, fill = score > 0)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +
  ylab("Average sentiment score")
```

## Looking by word

It's worth discovering *why* some topics ended up more positive then others. For that, we can examine the total positive and negative contributions of each word:

```{r}
contributions <- usenet_words %>%
  inner_join(AFINN, by = "word") %>%
  group_by(word) %>%
  summarize(occurences = n(),
            contribution = sum(score))

contributions
```

We can visualize which words had the most effect.

```{r}
contributions %>%
  top_n(25, abs(contribution)) %>%
  mutate(word = reorder(word, contribution)) %>%
  ggplot(aes(word, contribution, fill = contribution > 0)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip()
```

These words look generally reasonable as indicators of each message's sentiment, but we can spot possible problems with the approach. "True" could just as easily be a part of "not true" or a similar negative expression, and the words "God" and "Jesus" are apparently very common on Usenet but could easily be used in many contexts.

We may also care about which words contributed the most *within each board*. We can calculate each word's contribution to each board's sentiment score from our `word_board_sentiments` variable:

```{r top_sentiment_words, dependson = "word_board_sentiments", fig.height = 10, fig.width = 10}
top_sentiment_words <- word_board_sentiments %>%
  mutate(contribution = score * n / sum(n))

top_sentiment_words %>%
  group_by(board) %>%
  top_n(8, abs(contribution)) %>%
  ungroup() %>%
  mutate(board = reorder(board, contribution),
         word = reorder(word, contribution)) %>%
  ggplot(aes(word, contribution, fill = contribution > 0)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ board, scales = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

We can also see how much sentiment is confounded with topic in this particular approach. An atheism board is likely to discuss "god" in detail even in a negative context, and we can see it makes the board look more positive. Similarly, the negative contribution of the word "gun" to the "talk.politics.guns" board would occur even if the board were positive.

## Sentiment analysis by message

We can also try finding the most positive and negative *messages*:

```{r}
sentiment_messages <- usenet_words %>%
  inner_join(AFINN, by = "word") %>%
  group_by(board, id) %>%
  summarize(sentiment = mean(score),
            words = n()) %>%
  ungroup() %>%
  filter(words >= 5)
```

As a simple measure to reduce the role of randomness, we filtered out messages that had fewer than five words that contributed to sentiment.

What was the most positive messages?

```{r}
sentiment_messages %>%
  arrange(desc(sentiment))
```

Let's check this by looking at the message?

```{r print_message, dependson = "cleaned_text"}
print_message <- function(message_id) {
  cleaned_text %>%
    filter(id == message_id) %>%
    filter(text != "") %>%
    .$text %>%
    cat(sep = "\n")
}

print_message(53560)
```

Looks like it's because the message uses the word "winner" a lot! How about the most negative message? Turns out it's also from the hockey site, but has a very different attitude:

```{r dependson = "sentiment_messages"}
sentiment_messages %>%
  arrange(sentiment)

print_message(53907)
```



## N-grams

We can also 

```{r usenet_2grams}
usenet_digrams <- cleaned_text %>%
  unnest_tokens(digram, text, token = "ngrams", n = 2)
```

```{r dependson = "usenet_2grams"}
usenet_digram_counts <- usenet_digrams %>%
  count(board, digram)
```

```{r usenet_digram_counts}
digram_tf_idf <- usenet_digram_counts %>%
  bind_tf_idf(digram, board, n)
```

```{r dependson = "usenet_digram_counts"}
negate_words <- c("not", "without", "no", "isn't", "can't", "don't",
                  "won't", "couldn't")

usenet_digram_counts %>%
  ungroup() %>%
  separate(digram, c("word1", "word2"), sep = " ") %>%
  filter(word1 %in% negate_words) %>%
  count(word1, word2, wt = n, sort = TRUE) %>%
  inner_join(AFINN, by = c(word2 = "word")) %>%
  mutate(contribution = score * nn) %>%
  top_n(10, abs(contribution)) %>%
  ungroup() %>%
  mutate(word2 = reorder(word2, contribution)) %>%
  ggplot(aes(word2, contribution, fill = contribution > 0)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~ word1, scales = "free", nrow = 2) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

TODO: Something is wrong with that plot? Shouldn't each panel have 10 words? Missing a `group_by`?

```{r}
```


```{r eval = FALSE, echo = FALSE}
# we're not going to use this one
metadata <- raw_text %>%
  group_by(id) %>%
  filter(cumsum(text == "") == 0) %>%
  ungroup() %>%
  separate(text, c("header", "content"),
           sep = ": ", extra = "merge", fill = "right") %>%
  filter(!is.na(content)) %>%
  mutate(header = str_replace_all(str_to_lower(header), "-", "_")) %>%
  distinct(id, header, .keep_all = TRUE) %>%
  spread(header, content)
```
